### 2021-02-05 R/analysis.Rmd
initial commit

---
title: "Analysis"
author: "Bertrand Wilden"
date: "1/3/2021"
output: html_document
---

```{r}
library(tidyverse)
library(bper)
library(caret)
library(wru)
```

```{r}
census_api <- "5e4c2b8438222753a7f4753fa78855eca73b9950"

load(file = "data/census_fl.RData")
load(file = "data/census_nc.RData")
load_bperdata(destination = "data", include_blocks = T)

load(file = "data/nc_fl.rda")
load(file = "data/nc_fl_wru.rda")
```

```{r}
source(here("R", "funcs", "classify_and_report.R"))
nc_fl_wru %>% 
  sample_n(1000) %>% 
  classify_and_report(wru = T, wru_geo = "county", test_name = "dk")
```

```{r}
source(here("R", "funcs", "classify_and_report.R"))

# Set up test data
sample_vec <- sample(1:nrow(nc_fl), 5e6, replace = FALSE)
nc_fl_tests <- nc_fl[sample_vec, ] %>% 
  select(-c(birth_year, female))
nc_fl_wru_tests <- nc_fl_wru[sample_vec, ]

# Test results holder
tests <- data.frame()

# Full comparisons
wru_all <- nc_fl_wru_tests %>% 
  classify_and_report(wru = T, wru_geo = "block", test_name = "wru_all")
bper_all <- nc_fl_tests %>% 
  classify_and_report(test_name = "bper_all")
tests <- rbind(tests, wru_all, bper_all)

# Place
wru_place <- nc_fl_wru_tests %>% 
  filter(!is.na(place)) %>% 
  classify_and_report(wru = T, wru_geo = "place", test_name = "wru_place")
bper_place <- nc_fl_tests %>% 
  select(-c(block, apartment)) %>% 
  filter(!is.na(place)) %>% 
  classify_and_report(test_name = "bper_place")
tests <- rbind(tests, wru_place, bper_place)

# Zip
bper_zip <- nc_fl_tests %>% 
  select(-c(block, place, apartment)) %>% 
  classify_and_report(test_name = "bper_zip")
tests <- rbind(tests, bper_zip)

# County
wru_county <- nc_fl_wru_tests %>% 
  classify_and_report(wru = T, wru_geo = "county", test_name = "wru_county")
bper_county <- nc_fl_tests %>% 
  select(-c(block, place, zip, apartment)) %>% 
  classify_and_report(test_name = "bper_county")
tests <- rbind(tests, wru_county, bper_county)

# State
bper_state <- nc_fl_tests %>% 
  select(-c(block, place, zip, county, apartment)) %>% 
  classify_and_report(test_name = "bper_state")
tests <- rbind(tests, bper_state)

# No Party
wru_all_np <- nc_fl_wru_tests %>% 
  select(-PID) %>% 
  classify_and_report(wru = T, wru_geo = "block", wru_party = F, test_name = "wru_all_np")
bper_all_np <- nc_fl_tests %>% 
  select(-party) %>% 
  classify_and_report(test_name = "bper_all_np")
tests <- rbind(tests, wru_all_np, bper_all_np)

wru_place_np <- nc_fl_wru_tests %>% 
  select(-PID) %>%
  filter(!is.na(place)) %>% 
  classify_and_report(wru = T, wru_geo = "place", wru_party = F, test_name = "wru_place_np")
bper_place_np <- nc_fl_tests %>% 
  select(-c(party, block, apartment)) %>% 
  filter(!is.na(place)) %>% 
  classify_and_report(test_name = "bper_place_np")
tests <- rbind(tests, wru_place_np, bper_place_np)

bper_zip_np <- nc_fl_tests %>% 
  select(-c(party, block, place, apartment)) %>% 
  classify_and_report(test_name = "bper_zip_np")
tests <- rbind(tests, bper_zip_np)

wru_county_np <- nc_fl_wru_tests %>% 
  select(-PID) %>% 
  classify_and_report(wru = T, wru_geo = "county", wru_party = F, test_name = "wru_county_np")
bper_county_np <- nc_fl_tests %>% 
  select(-c(party, block, place, zip, apartment)) %>% 
  classify_and_report(test_name = "bper_county_np")
tests <- rbind(tests, wru_county_np, bper_county_np)

bper_state_np <- nc_fl_tests %>% 
  select(-c(party, block, place, zip, county, apartment)) %>% 
  classify_and_report(test_name = "bper_state_np")
tests <- rbind(tests, bper_state_np)

save(tests, file = here("data", "tests.rda"))
```