### 2021-04-22 S21_American_Workshop/S21_American_Workshop.Rmd
s21 american workshop presentation

---
title: "Improved Bayesian Ethnorace Prediction"
author: "Bertrand Wilden"
date: "`r format(Sys.time(), '%d %B, %Y')`"
fontsize: 10pt
output: 
  binb::metropolis:
    keep_tex: no
    pandoc_args: ["--lua-filter=hideslide.lua"]
urlcolor: black
linkcolor: black
header-includes:
  - \definecolor{shadecolor}{RGB}{230,230,230}
  # - \setbeamercolor{frametitle}{bg=black}
---

```{r, include=FALSE}
library(tidyverse)
library(ggthemes)
library(gridExtra)
library(here)
library(grid)
library(knitr)
opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

load(here("data", "tests.rda"))

good_color <- "palegreen4"
bad_color <- "slategrey"
```

## Motivation

::: incremental
- Race/ethnicity data not collected
- Aggregate proportions may be unreliable---ecological inference issues
- Existing studies using ethnorace prediction methods:
  - Protests and voting (Enos, Kaufman & Sands 2019)
  - Representation in local politics (Abott Magazinnik 2020)
  - Racial disparities in voter turnout (Fraga 2018), campaign finance (Grumbach & Sahm 2020; Grumbach, Sahn & Staszak 2020), and evictions (Hepburn, Louis & Desmond 2020)
- Prediction accuracy affects all downstream empirical results in these studies!
:::

# Method

## Method - overview

::: incremental
- Use Bayes' Rule to combine known ethnorace probabilities of attributes into posterior probabilities
- The ethnorace category with the highest posterior probability becomes the predicted ethnorace for the individual
:::

## Method - overview

**Outputs:**

- US Census categorizations
  - Asian American or Pacific Islander, American Indian and Alaska Native, Black/African American, Hispanic, Other/two or more races, White

## Method - overview

**Inputs:**

- First Names
- Surnames
- Geolocations (state, county, place, ZIP, block)
- Party ID
- Age
- Gender
- Multi-unit Occupancy

## Method - overview

Example:

- Given a profile of attributes: **First Name** BERT, **Surname** WILDEN, **ZIP Code** 92037
  - Asian = 0.03
  - Native American = 0.08
  - Hispanic = 0.004 
  - Black = 0.08
  - White = 0.53
  - Other Race = 0.28
- Prediction: *White*

## Method - details

Bayes' Rule

$$Pr(R=r|X) = \frac{Pr(X|R=r)Pr(R=r)}{Pr(X)} \qquad \qquad (1)$$
Where

- *R* = "true" ethnorace
- *r* = predicted ethnorace
- *X* = joint probability of having a particular profile of characteristics (first name, surname, geolocation, party ID, age, gender, and address type)

## Method - details

Assuming conditional independence of ethnorace and attributes, Equation (1) becomes:
$$
Pr(R=r|X) = \frac{Pr(R=r|x')\prod\limits_{j=1}^6Pr(x_j|R=r)}{\sum\limits_{i=1}^{6} Pr(R=r_i|x')\prod\limits_{j=1}^6Pr(x_j|R=r_i)} \qquad \qquad (2)
$$
Conditional independence:

- Knowing both a particular attribute of an individual, and that individual's ethnorace, should give us no extra knowledge of any other attribute for that individual
  - Probably violated, but the maximum a posteriori predictions remain accurate

# Validation

## Validation

::: incremental
- Validated using the North Carolina and Florida voter files
  - 2% Asian, 16% Black, 11% Hispanic, 6% Other Race, 65% White
  - $n = 21,164,503$
- Imputed ethnorace predictions compared with individuals' self-reported ethnorace
:::

## Validation - metrics

::: incremental
- **Accuracy**: proportion of sample correctly predicted
- Group specific:
  - **Precision**: proportion of correct predictions by group (i.e. what is the probability that an individual predicted to be White is White?)
  - **Recall**: proportion of group correctly predicted (i.e. what proportion of Whites were correctly classified?)
:::

## Validation - Accuracy

```{r, fig.cap="Accuracy Scores by Input Data"}
acc_bper <- tests %>% 
  filter(test_method == "bper") %>% 
  select(test_type, accuracy)

ggplot(acc_bper, aes(x = test_type, y = accuracy, group = 1)) +
  geom_line(color = good_color) +
  geom_point(size = 3, color = good_color) +
  theme_minimal() +
  ylim(0.75, 0.85) +
  labs(y = "Accuracy Score") +
  scale_x_discrete(name = "Model Inputs", limits = rev(c("all", "all_np", "zip", "zip_np", "place", "place_np", "county", "county_np", "state", "state_np")),
                   labels = rev(c(
                     "all" = "Block \n Multi-Unit \n Party ID \n First Name \n Last Name", 
                     "place" = "Place \n Party ID \n First Name \n Last Name", 
                     "county" = "County \n Party ID \n First Name \n Last Name", 
                     "all_np" = "Block \n Multi-Unit \n First Name \n Last Name", 
                     "place_np" = "Place \n First Name \n Last Name", 
                     "county_np" = "County \n First Name \n Last Name",
                     "zip" = "ZIP \n Party ID \n First Name \n Last Name",
                     "zip_np" = "ZIP \n First Name \n Last Name",
                     "state" = "State \n Party ID \n First Name \n Last Name",
                     "state_np" = "State \n First Name \n Last Name"
                   ))) +
  theme(text = element_text(size = 16), axis.text.x = element_text(size = 10),
        axis.title.y = element_text(margin = margin(r = 20)),
        panel.grid.major.y = element_line(color = "darkgrey"))
```

## Validation - Precision and Recall

```{r, fig.cap="Precision/Recall Scores by Ethnorace and Input Data"}
metrics_bper <- tests %>% 
  filter(test_method == "bper") %>% 
  select(test_type, contains(c("prec", "rec"))) %>% 
  pivot_longer(cols = contains(c("prec", "rec"))) %>% 
  separate(name, c("Metric", "Race"), sep = "_") %>% 
  filter(Race != "other") %>% 
  mutate(Metric = case_when(Metric == "prec" ~ "Precision",
                            Metric == "rec" ~ "Recall"),
         Race = case_when(Race == "api" ~ "Asian",
                          Race == "black" ~ "Black",
                          Race == "hispanic" ~ "Hispanic",
                          Race == "white" ~ "White"))

ggplot(metrics_bper, aes(x = test_type, y = value, group = Metric)) +
  geom_line(aes(color = Metric), size = 0.3) +
  geom_point(aes(color = Metric), size = 2) +
  theme_minimal() +
  labs(x = "", y = "Score") +
  scale_color_manual(values = c("gold3", "skyblue4")) +
  scale_x_discrete(name = "Model Inputs", limits = rev(c("all", "all_np", "zip", "zip_np", "place", "place_np", "county", "county_np", "state", "state_np")),
                   labels = rev(c(
                     "all" = "Block", 
                     "place" = "Place", 
                     "county" = "County ", 
                     "all_np" = "Block (np)", 
                     "place_np" = "Place (np)", 
                     "county_np" = "County (np)",
                     "zip" = "ZIP",
                     "zip_np" = "ZIP (np)",
                     "state" = "State",
                     "state_np" = "State (np)"
                   ))) +
  facet_wrap(~ Race) +
  theme(axis.text.x = element_text(angle = 50), text = element_text(size = 16), 
        axis.title.y = element_text(margin = margin(r = 20)),
        panel.grid.major.y = element_line(color = "darkgrey"))
```

# Comparison to Existing Methods

## Comparison With Existing Methods

::: incremental
- Imai and Khanna (2016), R package **`wru`**
- R package **`bper`**[^1]
  - Better predictions
    - First names and address type as additional inputs
    - "Smoothing" for missing values
    - Averages over posterior probabilities for each choice of "prior"
  - More flexibility
    - State and ZIP Code options
    - American Indian and Alaska Native predictions
    - Faster/easier implementation
:::

[^1]: Bayesian Prediction for Ethnicity and Race. https://github.com/bwilden/bper

## Comparison - Accuracy

```{r, fig.cap="Accuracy Score Comparison to `wru`"}
plot_tests <- tests %>% 
  filter(!(test_type %in% c("zip", "zip_np", "state", "state_np")))

accuracy <- plot_tests %>% 
  select(test_method, test_type, accuracy) %>% 
  group_by(test_type) %>% 
  summarise(diff = last(accuracy) - first(accuracy), bper_val = last(accuracy))

ggplot(accuracy, aes(x = test_type, y = diff)) +
  geom_bar(stat = "identity", fill = good_color) +
  theme_minimal() + 
  labs(y = "Difference in Accuracy", x = "") +
  scale_x_discrete(name = "Model Inputs", limits = rev(c("all", "all_np", "place", "place_np", "county", "county_np")),
                   labels = rev(c(
                     "all" = "Block \n Multi-Unit \n Party ID \n First Name \n Last Name", 
                     "place" = "Place \n Party ID \n First Name \n Last Name", 
                     "county" = "County \n Party ID \n First Name \n Last Name", 
                     "all_np" = "Block \n Multi-Unit \n First Name \n Last Name", 
                     "place_np" = "Place \n First Name \n Last Name", 
                     "county_np" = "County \n First Name \n Last Name"))) +
  theme(text = element_text(size = 16), axis.title.y = element_text(margin = margin(r = 20)),
        panel.grid.major.y = element_line(color = "darkgrey", linetype = "dashed"))
```

## Comparison - Precision/Recall Asian

```{r, fig.cap="Precision and Recall Score Comparison to `wru`: Asian"}
api <- plot_tests %>% 
  select(test_method, test_type, prec_api, rec_api) %>% 
  group_by(test_type) %>% 
  summarise(prec_diff = last(prec_api) - first(prec_api), prec_bper = last(prec_api),
            rec_diff = last(rec_api) - first(rec_api), rec_bper = last(rec_api)) %>% 
  ungroup() %>% 
  mutate(prec_color = if_else(prec_diff < 0, bad_color, good_color),
         rec_color = if_else(rec_diff < 0, bad_color, good_color))

prec_api <-
  ggplot(api, aes(x = test_type, y = prec_diff, fill = prec_color)) +
  geom_bar(stat = "identity") +
  scale_fill_identity(guide = F) +
  theme_minimal() + 
  ylim(0, 0.145) +
  labs(y = "Difference in Precision", x = "", title = "") +
  scale_x_discrete(name = "", limits = rev(c("all", "all_np", "place", "place_np", "county", "county_np")),
                   labels = rev(c(
                     "all" = "", 
                     "place" = "", 
                     "county" = "", 
                     "all_np" = "", 
                     "place_np" = "", 
                     "county_np" = ""))
                   ) +
  theme(text = element_text(size = 16), axis.title.y = element_text(margin = margin(r = 20)),
        panel.grid.major.y = element_line(color = "darkgrey", linetype = "dashed"))

rec_api <-
 ggplot(api, aes(x = test_type, y = rec_diff, fill = rec_color)) +
  geom_bar(stat = "identity") +
  scale_fill_identity(guide = F) +
  theme_minimal() + 
  ylim(-0.14, 0.05) +
  labs(y = "Difference in Recall", x = "") +
  scale_x_discrete(name = "Model Inputs", limits = rev(c("all", "all_np", "place", "place_np", "county", "county_np")),
                   labels = rev(c(
                     "all" = "Block \n Multi-Unit \n Party ID \n First Name \n Last Name", 
                     "place" = "Place \n Party ID \n First Name \n Last Name", 
                     "county" = "County \n Party ID \n First Name \n Last Name", 
                     "all_np" = "Block \n Multi-Unit \n First Name \n Last Name", 
                     "place_np" = "Place \n First Name \n Last Name", 
                     "county_np" = "County \n First Name \n Last Name"))) +
  theme(text = element_text(size = 16), axis.title.y = element_text(margin = margin(r = 20)),
        panel.grid.major.y = element_line(color = "darkgrey", linetype = "dashed"))

grid.arrange(prec_api, rec_api, nrow = 2)
```

## Comparison - Precision/Recall Black

```{r, fig.cap="Precision and Recall Score Comparison to `wru`: Black"}
black <- plot_tests %>% 
  select(test_method, test_type, prec_black, rec_black) %>% 
  group_by(test_type) %>% 
  summarise(prec_diff = last(prec_black) - first(prec_black), prec_bper = last(prec_black),
            rec_diff = last(rec_black) - first(rec_black), rec_bper = last(rec_black)) %>% 
  ungroup() %>% 
  mutate(prec_color = if_else(prec_diff < 0, bad_color, good_color),
         rec_color = if_else(rec_diff < 0, bad_color, good_color))

prec_black <-
  ggplot(black, aes(x = test_type, y = prec_diff, fill = prec_color)) +
  geom_bar(stat = "identity") +
  scale_fill_identity(guide = F) +
  theme_minimal() + 
  ylim(0, 0.3) +
  labs(y = "Difference in Precision", x = "", title = "") +
  scale_x_discrete(name = "", limits = rev(c("all", "all_np", "place", "place_np", "county", "county_np")),
                   labels = rev(c(
                     "all" = "", 
                     "place" = "", 
                     "county" = "", 
                     "all_np" = "", 
                     "place_np" = "", 
                     "county_np" = ""))
                   ) +
  theme(text = element_text(size = 16), axis.title.y = element_text(margin = margin(r = 20)),
        panel.grid.major.y = element_line(color = "darkgrey", linetype = "dashed"))

rec_black <-
 ggplot(black, aes(x = test_type, y = rec_diff, fill = rec_color)) +
  geom_bar(stat = "identity") +
  scale_fill_identity(guide = F) +
  theme_minimal() + 
  ylim(-0.3, 0.05) +
  labs(y = "Difference in Recall", x = "") +
  scale_x_discrete(name = "Model Inputs", limits = rev(c("all", "all_np", "place", "place_np", "county", "county_np")),
                   labels = rev(c(
                     "all" = "Block \n Multi-Unit \n Party ID \n First Name \n Last Name", 
                     "place" = "Place \n Party ID \n First Name \n Last Name", 
                     "county" = "County \n Party ID \n First Name \n Last Name", 
                     "all_np" = "Block \n Multi-Unit \n First Name \n Last Name", 
                     "place_np" = "Place \n First Name \n Last Name", 
                     "county_np" = "County \n First Name \n Last Name"))) +
  theme(text = element_text(size = 16), axis.title.y = element_text(margin = margin(r = 20)),
        panel.grid.major.y = element_line(color = "darkgrey", linetype = "dashed"))

grid.arrange(prec_black, rec_black, nrow = 2)
```

## Comparison - Precision/Recall Hispanic

```{r, fig.cap="Precision and Recall Score Comparison to `wru`: Hispanic"}
hispanic <- plot_tests %>% 
  select(test_method, test_type, prec_hispanic, rec_hispanic) %>% 
  group_by(test_type) %>% 
  summarise(prec_diff = last(prec_hispanic) - first(prec_hispanic), prec_bper = last(prec_hispanic),
            rec_diff = last(rec_hispanic) - first(rec_hispanic), rec_bper = last(rec_hispanic)) %>% 
  ungroup() %>% 
  mutate(prec_color = if_else(prec_diff < 0, bad_color, good_color),
         rec_color = if_else(rec_diff < 0, bad_color, good_color))

prec_hispanic <-
  ggplot(hispanic, aes(x = test_type, y = prec_diff, fill = prec_color)) +
  geom_bar(stat = "identity") +
  scale_fill_identity(guide = F) +
  theme_minimal() + 
  ylim(0, 0.04) +
  labs(y = "Difference in Precision", x = "", title = "") +
  scale_x_discrete(name = "", limits = rev(c("all", "all_np", "place", "place_np", "county", "county_np")),
                   labels = rev(c(
                     "all" = "", 
                     "place" = "", 
                     "county" = "", 
                     "all_np" = "", 
                     "place_np" = "", 
                     "county_np" = ""))
                   ) +
  theme(text = element_text(size = 16), axis.title.y = element_text(margin = margin(r = 20)),
        panel.grid.major.y = element_line(color = "darkgrey", linetype = "dashed"))

rec_hispanic <-
 ggplot(hispanic, aes(x = test_type, y = rec_diff, fill = rec_color)) +
  geom_bar(stat = "identity") +
  scale_fill_identity(guide = F) +
  theme_minimal() + 
  ylim(0, 0.04) +
  labs(y = "Difference in Recall", x = "") +
  scale_x_discrete(name = "Model Inputs", limits = rev(c("all", "all_np", "place", "place_np", "county", "county_np")),
                   labels = rev(c(
                     "all" = "Block \n Multi-Unit \n Party ID \n First Name \n Last Name", 
                     "place" = "Place \n Party ID \n First Name \n Last Name", 
                     "county" = "County \n Party ID \n First Name \n Last Name", 
                     "all_np" = "Block \n Multi-Unit \n First Name \n Last Name", 
                     "place_np" = "Place \n First Name \n Last Name", 
                     "county_np" = "County \n First Name \n Last Name"))) +
  theme(text = element_text(size = 16), axis.title.y = element_text(margin = margin(r = 20)),
        panel.grid.major.y = element_line(color = "darkgrey", linetype = "dashed"))

grid.arrange(prec_hispanic, rec_hispanic, nrow = 2)
```

## Comparison - Precision/Recall White

```{r, fig.cap="Precision and Recall Score Comparison to `wru`: White"}
white <- plot_tests %>% 
  select(test_method, test_type, prec_white, rec_white) %>% 
  group_by(test_type) %>% 
  summarise(prec_diff = last(prec_white) - first(prec_white), prec_bper = last(prec_white),
            rec_diff = last(rec_white) - first(rec_white), rec_bper = last(rec_white)) %>% 
  ungroup() %>% 
  mutate(prec_color = if_else(prec_diff < 0, bad_color, good_color),
         rec_color = if_else(rec_diff < 0, bad_color, good_color))

prec_white <-
  ggplot(white, aes(x = test_type, y = prec_diff, fill = prec_color)) +
  geom_bar(stat = "identity") +
  scale_fill_identity(guide = F) +
  theme_minimal() + 
  ylim(-0.06, 0.02) +
  labs(y = "Difference in Precision", x = "", title = "") +
  scale_x_discrete(name = "", limits = rev(c("all", "all_np", "place", "place_np", "county", "county_np")),
                   labels = rev(c(
                     "all" = "", 
                     "place" = "", 
                     "county" = "", 
                     "all_np" = "", 
                     "place_np" = "", 
                     "county_np" = ""))
                   ) +
  theme(text = element_text(size = 16), axis.title.y = element_text(margin = margin(r = 20)),
        panel.grid.major.y = element_line(color = "darkgrey", linetype = "dashed"))

rec_white <-
 ggplot(white, aes(x = test_type, y = rec_diff, fill = rec_color)) +
  geom_bar(stat = "identity") +
  scale_fill_identity(guide = F) +
  theme_minimal() + 
  ylim(0, 0.08) +
  labs(y = "Difference in Recall", x = "") +
  scale_x_discrete(name = "Model Inputs", limits = rev(c("all", "all_np", "place", "place_np", "county", "county_np")),
                   labels = rev(c(
                     "all" = "Block \n Multi-Unit \n Party ID \n First Name \n Last Name", 
                     "place" = "Place \n Party ID \n First Name \n Last Name", 
                     "county" = "County \n Party ID \n First Name \n Last Name", 
                     "all_np" = "Block \n Multi-Unit \n First Name \n Last Name", 
                     "place_np" = "Place \n First Name \n Last Name", 
                     "county_np" = "County \n First Name \n Last Name"))) +
  theme(text = element_text(size = 16), axis.title.y = element_text(margin = margin(r = 20)),
        panel.grid.major.y = element_line(color = "darkgrey", linetype = "dashed"))

grid.arrange(prec_white, rec_white, nrow = 2)
```

## Next Steps

::: incremental
- Larger research agenda on ethics in AI/machine learning for the study of race
  - Where do mis-predictions happen?
- Replicate existing studies which use the old method
:::

---

Thank you!

bwilden@ucsd.edu